\documentclass{llncs}
\usepackage{etex}

\usepackage{xcolor}
\usepackage{enumitem,amsmath,amssymb}
\usepackage{url}    % used for \url and \burl
\usepackage[linesnumbered,boxed,noline,noend]{algorithm2e}
\def\defaultHypSeparation{\hskip.1in}

\usepackage{tikz}
\usepackage{subfig}
\usepackage{array,booktabs,multirow}
\usepackage{placeins}

\usepackage{logictools}
\usepackage{prooftheory}
\usepackage{comment}
\usepackage{mathenvironments}
\usepackage{drawproof}
\usepackage{bussproofs}
\usepackage{tensor}
\usepackage{mathtools}
\usepackage{amsmath}

\usepackage{graphicx}
%\usepackage{caption}
%\usepackage{subcaption}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}


\newcommand{\freevar}[1]{\mathrm{FV}(#1)}

\newcommand{\Vertices}[1]{V_{#1}}
\newcommand{\Edges}[1]{E_{#1}}
\newcommand{\Conclusion}[1]{\clause_{#1}}

\newcommand{\axiom}[1]{\widehat{#1}}
\newcommand{\n}{v}
\newcommand{\raiz}[1]{\rho(#1)}

\newcommand{\pedge}[3]{\ensuremath{\raiz{#1} \xrightarrow{#2} \raiz{#3}}}


\newcommand\inlineeqno{\stepcounter{equation}\ (\theequation)}


% Contraction
\newcommand{\con}[3]{\lfloor #1 \rfloor_{#2}^{#3}}

% Resolution
%\newcommand{\res}[6]{#1 \tensor[^{#2}_{#3}]{\odot}{^{#4}_{#5}} #6}
%\newcommand{\res}[6]{#1 \prescript{#2}{#3}{\odot^{#4}_{#5}} #6}

\newcommand{\res}[4]{\mathrel{\operatorname*{\odot}_{#1 #3}^{#2 #4}}}

\title{Towards the Compression of First-Order Resolution Proofs by Lowering Unit Clauses}

\author{
  Jan Gorzny\inst{1}
  \thanks{Supported by the Google Summer of Code 2014 program.}
  \and 
  Bruno Woltzenlogel Paleo\inst{2}
  \thanks{Stipendiat der \"Osterreichischen Akademie der Wissenschaften (APART).}
}

\authorrunning{J.\~Gorzny \and B.\~Woltzenlogel Paleo}

\institute{
  \email{jgorzny@uvic.ca}, University of Victoria, Canada
  \and 
  \email{bruno@logic.at}, Vienna University of Technology, Austria
}




\begin{document}

\maketitle


\begin{abstract}
The recently developed {\LowerUnits} algorithm compresses
propositional resolution proofs generated by SAT- and SMT-solvers by postponing and lowering resolution inferences involving unit clauses, which have exactly one literal. This paper describes a generalization of this algorithm to the case of first-order resolution proofs generated by automated theorem provers. An empirical evaluation of a simplified version of this algorithm on hundreds of proofs shows promising results.
\end{abstract}


\setcounter{footnote}{0}

\section{Introduction}

Most of the effort in automated reasoning so far has been dedicated to the design and implementation of proof systems and efficient theorem proving procedures. As a result, saturation-based first-order automated theorem provers have achieved a high degree of maturity, with resolution and superposition being among the most common underlying proof calculi. Proof production is an essential feature of modern state-of-the-art provers and proofs are crucial for applications where the user requires certification of the answer provided by the prover. Nevertheless, efficient proof production is 
non-trivial,
%non-trivial \cite{SchultzAPPA}, 
and it is to be expected that the best, most efficient, provers do not necessarily generate the best, least redundant, proofs. Therefore, it is a timely moment to develop methods that post-process and simplify proofs. While the foundational problem of simplicity of proofs can be traced back at least to Hilbert's 24th Problem, the maturity of automated deduction has made it particularly relevant today.  

For proofs generated by SAT- and SMT-solvers, which use propositional resolution as the basis for the DPLL and CDCL decision procedures, there is now a wide variety of proof compression techniques. Algebraic properties of the resolution
operation that might be useful for compression were investigated in \cite{bwp10}.
%Compression algorithms based on rearranging and sharing chains of resolution inferences have been
%developed in \cite{Amjad07} and \cite{Sinz}. 
% Cotton \cite{CottonSplit} proposed an algorithm that
%compresses a refutation by repeteadly splitting it into a proof of a heuristically chosen literal $\ell$
%and a proof of $\dual{\ell}$, and then resolving them to form a new refutation.  
The {\ReduceReconstruct} algorithm \cite{RedRec} searches for locally redundant
subproofs that can be rewritten into subproofs of stronger clauses and with fewer resolution steps.
A linear time proof compression algorithm based on partial
regularization was proposed in \cite{RP08} and improved in \cite{LURPI}. Furthermore, \cite{LURPI} described a linear time algorithm called {\LowerUnits} that delays resolution with unit clauses.

In contrast, for first-order theorem provers, there has been up to now (to the best of our knowledge) no attempt to design and implement an algorithm capable of taking a first-order resolution DAG-proof and efficiently simplifying it, outputting a possibly shorter pure first-order resolution DAG-proof. There are algorithms aimed at simplifying first-order sequent calculus tree-like proofs, based on cut-introduction and Herbrand sequents \cite{BrunoLPAR,Hetzl,HerbrandSequent}.
%, and while in principle resolution DAG-proofs can be translated to sequent-calculus tree-like proofs (and then back), such translations lead to undesirable efficiency overheads. 
There is also an algorithm \cite{LPARCzech} that looks for terms that occur often in any TSTP \cite{TPTP} proof (including first-order resolution DAG-proofs) and introduces abbreviations for these terms. However, as the definitions of the abbreviations are not part of the output proof, it cannot be checked by a pure first-order resolution proof checker.

In this paper, we initiate the process of lifting propositional proof compression techniques to the first-order case, starting with the simplest known algorithm: {\LowerUnits} (described in \cite{LURPI}). 
As shown in Section \ref{sec:Challenges}, even for this simple algorithm, the fact that first-order resolution makes use of unification leads to many challenges that simply do not exist in the propositional case. 
%In Section \ref{sec:FOLU} we describe an algorithm that overcomes these challenges. As this algorithm has quadratic run-time complexity with respect to the input proof length, in Section \ref{sec:SimpleFOLU} we describe a variation of this algorithm, which has linear run-time complexity and is easier to implement, at the cost of compressing less. 
In Section \ref{sec:SimpleFOLU} we describe an easy to implement algorithm with  linear time complexity (with respect to the proof length) which partially overcomes these challenges. 
In Section \ref{sec:exp} we present experimental results obtained by applying this algorithm on hundreds of proofs generated with the{\
%SPASS} theorem prover \cite{SPASS}. 
SPASS} theorem prover. 
The next section introduces the first-order resolution calculus using notations that are more convenient for describing proof transformation operations.






\input{sec-Resolution}
\input{sec-Challenges}
\input{sec-SimpleFOLU}
\input{sec-Exp}



\section{Conclusions and Future Work}

{\SFOLowerUnits} is our first attempt to lift a propositional proof compression algorithm to the first-order case. We consider this algorithm a prototype, useful to evaluate whether this approach is promising. The experimental results discussed in the previous section are encouraging, especially in comparison with existing results for the propositional case. In the near future, we shall seek improvements of this algorithm as well as other ways to overcome the difficulties related to the post-deletion unifiability property.

The difficulties related to unit reintroduction suggest that other propositional proof compression algorithms that do not require reintroduction (e.g. {\RecyclePivotsIntersection} \cite{LURPI}) might need less sophisticated bookkeeping when lifted to first-order.

The efficiency and versatility of contemporary automated theorem provers depend on inference rules (e.g. equality rules) and techniques (e.g. splitting) that go beyond the pure resolution calculus. The eventual generalization of the compression algorithms to support such extended calculi will be essential for their usability on a wider range of problems.

%\vspace{-10pt}
%\paragraph{Acknowledgments:}


\begin{footnotesize}
%\bibliographystyle{splncs}
\bibliographystyle{plain}
\bibliography{biblio}
\end{footnotesize}


\appendix

%\input{sec-PropositionalLU}
%\input{sec-FOLU}



\end{document}

% vim: tw=100
