\section{The Resolution Calculus}

We assume that there are infinitely many variable symbols (e.g. $X$, $Y$, $Z$, $X_1$, $X_2$, \ldots), constant symbols (e.g. $a$, $b$, $c$, $a_1$, $a_2$, \ldots), function symbols of every arity (e.g $f$, $g$, $f_1$, $f_2$, \ldots) and predicate symbols of every arity (e.g. $p$, $q$, $p_1$, $p_2$,\ldots). A \emph{term} is any variable, constant or the application of an $n$-ary function symbol to $n$ terms.
An \emph{atomic formula} (\emph{atom}) is the application of an $n$-ary predicate symbol to $n$ terms. A \emph{literal} is an atom or the negation of an atom. The
\emph{complement} of a literal $\ell$ is denoted $\dual{\ell}$ (i.e. for any atom $p$,
$\dual{p} = \neg p$ and $\dual{\neg p} = p$). The set of all literals is denoted $\mathcal{L}$. A
\emph{clause} is a multiset of literals. $\bot$ denotes the \emph{empty clause}. A \emph{unit clause} is clause with a single literal. Sequent notation is used for clauses (i.e. $p_1,\ldots,p_n \seq q_1,\ldots, q_m$ denotes the clause $\{ \neg p_1,\ldots, \neg p_n, q_1, \ldots, q_m \}$).
$\freevar{t}$ (resp. $\freevar{\ell}$, $\freevar{\clause}$) denotes the set of variables in the term $t$ (resp. in the literal $\ell$ and in the clause $\clause$).
A \emph{substitution} $\{ X_1\backslash t_1, X_2 \backslash t_2, \ldots \}$ is a mapping from variables $\{ X_1, X_2, \ldots \}$ to, respectively, terms $\{t_1, t_2, \ldots \}$. The application of a substitution $\sigma$ to a term $t$, a literal $\ell$ or a clause $\clause$ results in, respectively, the term $t \sigma$, the literal $\ell \sigma$ or the clause $\clause \sigma$, obtained from $t$, $\ell$ and $\clause$ by replacing all occurrences of the variables in $\sigma$ by the corresponding terms in $\sigma$. The set of all substitutions is denoted $\mathcal{S}$. A \emph{unifier} of a set of literals is a substitution that makes all literals in the set equal.
A \emph{resolution proof} is a directed acyclic graph of clauses where the edges correspond to the inference rules of resolution and contraction (as explained in detail in Definition \ref{def:proof}). A \emph{resolution refutation} is a resolution proof with root $\bot$.


\begin{definition}[First-Order Resolution Proof] 
\label{def:proof} \hfill \\
A directed acyclic graph $\langle V, E, \clause \rangle$, where $V$ is a set of nodes and $E$ is a
set of edges labeled by literals and substitutions (i.e. $E \subset V \times \mathcal{L} \times \mathcal{S} \times V$ and $\n_1
\xrightarrow[\sigma]{\ell} \n_2$ denotes an edge from node $\n_1$ to node $\n_2$ labeled by the literal $\ell$ and the substitution $\sigma$), is a
proof of a clause $\clause$ iff it is inductively constructible according to the following cases:
%
\begin{itemize}
  \item \textbf{Axiom:} If $\Gamma$ is a clause, $\axiom{\Gamma}$ denotes some proof $\langle \{ \n \}, \varnothing,
    \Gamma \rangle$, where $\n$ is a new (axiom) node.
  \item \textbf{Resolution:} If $\psi_L$ is a proof $\langle V_L, E_L, \clause_L \rangle$ with $\ell_L \in \clause_L$ and
    $\psi_R$ is a proof $\langle V_R, E_R, \clause_R \rangle$ with $\ell_R \in \clause_R$, and 
    $\sigma_L$ and $\sigma_R$ are substitutions such that
    $\ell_L \sigma_L = \dual{\ell_R} \sigma_R$ and
    $\freevar{\left( \clause_L \setminus \left\{ \ell_L \right\} \right) \sigma_L} \cap 
     \freevar{\left( \clause_R
                    \setminus \left\{ \ell_R \right\} \right) \sigma_R} = \emptyset$, 
    then
    $\psi_L \res{\ell_L}{\sigma_L}{\ell_R}{\sigma_R} \psi_R$ denotes a proof $\langle V, E, \Gamma \rangle$ s.t.
    \begin{align*}
      V &= V_L \cup V_R \cup \{\n \} \\
      E &= E_L \cup E_R \cup
                    \left\{ \raiz{\psi_L} \xrightarrow[\sigma_L]{\ell_L} \n, 
                            \raiz{\psi_R} \xrightarrow[\sigma_R]{\ell_R} \n \right\} \\
     \Gamma &= \left( \clause_L \setminus \left\{ \ell_L \right\} \right) \sigma_L \cup \left( \clause_R
                    \setminus \left\{ \ell_R \right\} \right) \sigma_R 
    \end{align*}
    where $\n$ is a new (resolution) node and $\raiz{\varphi}$ denotes the root node of $\varphi$.
  \item \textbf{Contraction:} If $\psi'$ is a proof $\langle V', E', \clause' \rangle$ and $\sigma$ is a unifier of $\{\ell_1, \ldots \ell_n\}$ with $\{\ell_1, \ldots \ell_n\} \subseteq \clause'$, then, letting $\ell = \ell_k \sigma$ (for any $k \in \{1,\ldots, n\}$), $\con{\psi}{\sigma}{\ell}$ denotes a proof $\langle V, E, \Gamma \rangle$ s.t.
    \begin{align*}
      V &= V' \cup \{\n \} \\
      E &= E' \cup \{ \raiz{\psi'} \xrightarrow[\sigma]{\ell} \n \} \\
     \Gamma &= (\clause' \setminus \{ \ell_1, \ldots \ell_n \} ) \sigma \cup \{ \ell \}
    \end{align*}
    where $\n$ is a new (contraction) node and $\raiz{\varphi}$ denotes the root node of $\varphi$.
  \qed
\end{itemize}
\end{definition}




\noindent
If $\psi = \varphi_L \odot_{\ell} \varphi_R$, then $\varphi_L$ and $\varphi_R$ are \emph{direct
subproofs} of $\psi$ and $\psi$ is a \emph{child} of both $\varphi_L$ and $\varphi_R$. The
transitive closure of the direct subproof relation is the \emph{subproof} relation. A subproof which
has no direct subproof is an \emph{axiom} of the proof.
%
$\Vertices{\psi}$, $\Edges{\psi}$ and $\Conclusion{\psi}$
denote, respectively, the nodes, edges and proved clause (conclusion) of $\psi$.
