\section{Congruence Closure}
\label{sec:congruencedef}

In this section we define the concepts of congruence- relation and closure, which is the foundation of our proof compression method.
To this end we define terms and equations, which are notions that will be used throughout this chapter.
We close this section by proving some elementary properties of congruence relations.

Our definition of terms corresponds to what is usually called ground term in the context of first order logic.
Ground terms are terms that contain no first order logic variables.
Since we do not investigate non ground terms, we omit the complement and simply speak of terms.

\begin{definition}[Terms and Subterms]

Let $\mathcal{F}$ be a finite set of function symbols and $arity: \mathcal{F} \rightarrow \mathbb{N}$.
A tuple $\Sigma = \langle \mathcal{F}, arity \rangle$ is a \emph{signature}.
A function symbol with arity zero is a \emph{constant}, one with arity one is a \emph{unary} function symbol and one with arity two is \emph{binary}.
For a given signature $\Sigma$, the set of \emph{terms} $\mathcal{T}^{\Sigma}$ is defined inductively.

\begin{align*}
	\mathcal{T}^{\Sigma}_0 &= \{a \in \mathcal{F} \mid arity(a) = 0\}\\
	\mathcal{T}^{\Sigma}_{i+1} &= \{g(t_1,\ldots,t_n) \mid arity(g) = n \text{ and } t_1, \ldots, t_n \in \mathcal{T}_{i}\} \\
	\mathcal{T}^{\Sigma} &= \bigcup_{i\in \mathbb{N}} \mathcal{T}^{\Sigma}_{i}
\end{align*}

\noindent Let $g(t_1,\ldots,t_n) \in \mathcal{T}^{\Sigma}$, then $t_1,\ldots,t_n$ are \emph{direct subterms} of $g(t_1,\ldots,t_n)$.
The \emph{subterm} relation is the reflexive, transitive closure of the direct subterm relation.
A term of the form $g(t_1,\ldots,t_n)$ is a \emph{compound term}.

\end{definition}

\noindent Should $\Sigma$ be clear from context or of no relevance, we will omit it and write $\mathcal{T}$ instead of $\mathcal{T}^{\Sigma}$.

\begin{definition}[Equation]

Let $\mathcal{T}$ be a set of terms.
An \emph{equation} of $\mathcal{T}$ is a tuple of terms, i.e. an element of $\mathcal{T} \times \mathcal{T}$.

\end{definition}

\noindent For a set of equations $E$ we denote by $\mathcal{T}_E$ the set of terms used in $E$.
$$\mathcal{T}_E := \{t \mid t \text{ is subterm of some } u \text{, such that for some } v: (u,v) \in E \text{ or } (v,u) \in E\}$$

\begin{definition}[Congruence Relation]

Let $\mathcal{T}$ be a set of terms.
A relation $R \subseteq \mathcal{T} \times \mathcal{T}$ is a congruence relation, if has the following four properties:
\begin{itemize}
	\item reflexivity: for all $t \in \mathcal{T}: (t,t) \in R$
	\item symmetry: $(s,t) \in R$ then $(t,s) \in R$
	\item transitivity: $(r,s) \in R$ and $(s,t) \in R$ then $(r,t) \in R$
	\item compatibility: $g$ is a n-ary function symbol and for all $i = 1,\ldots,n$ $(t_i,s_i) \in R$ then $(g(t_1,\ldots,t_n),g(s_1,\ldots,s_n)) \in R$
\end{itemize}

\end{definition}

Clearly every congruence relation is also an equivalence relation (which is a reflexive, transitive and symmetric relation).
Therefore every congruence relation partitions its underlying set of terms $\mathcal{T}$ into congruence classes, such that two terms $(s,t)$ belong to the same class if and only if $(s,t) \in R$.
The relations $\emptyset$ and $\mathcal{T} \times \mathcal{T}$ are trivial congruence relations.

In this work we are interested in congruence relations induced by sets of equations.
In other words, we compute the partitioning of the terms such that two terms in the same partition are proven to be equal by the input set of equations.
To this end we define the notion of congruence closure of a set of equations.

\begin{definition}[Congruence Closure]

Let $E$ be a set of equations.
The set $E^* \supseteq E$ is the \emph{congruence closure} of $E$, 
if $E^*$ is a congruence relation on $\mathcal{T}_E$ and for every congruence relation $C$, such that $C \supseteq E$ follows $C \supseteq E^*$.
We write $E \models s \thickapprox t$ if $(s,t) \in E^*$ and say that $E$ is an \emph{explanation} for $s \thickapprox t$.
We call a pair $(s,t)$ in a congruence closure an \emph{equality} and we call an equality of compound terms $(g(t_1,\ldots,t_n),g(s_1,\ldots,s_n))$ such that for all $i = 1,\ldots,n$: $E \models t_i \thickapprox s_i$ a \emph{deduced equality}.
For a term $t \in \mathcal{T}_E$ the set of congruent terms $\{s \in \mathcal{T}_E \mid E \models s \thickapprox t\}$ is the \emph{congruence class} of $t$.

\end{definition}

It is easily seen that congruence relations are closed under intersection.
Therefore $E^*$ always exists.

\begin{proposition}[Properties of the $\models$ relation]
\label{prop:models}
The $\models$ relation is monotone: $E_1 \subseteq E_2$ and $E_1 \models s \thickapprox t$ implies $E_2 \models s \thickapprox t$ and consistent: $E \models s \thickapprox t$ and $E \cup \{(s,t)\} \models u \thickapprox v$ implies $E \models u \thickapprox v$. 

\end{proposition}

\begin{proof}

Monotonicity follows from the fact that congruence closure of $E_1$ is contained in the congruence closure of $E_2$.

Since the congruence closure of $E^*$ is $E^*$ itself, it follows that $E \models u \thickapprox v$ if and only if $E^* \models u \thickapprox v$.
Since $(s,t) \in E^*$, clearly it is the case that $E^* = (E \cup \{(s,t)\})^*$.
Therefore $E \cup \{(s,t)\} \models u \thickapprox v$ implies $(u,v) \in E^*$, i.e. $E \models u \thickapprox v$ or in other words, the $\models$ relation is consistent.

\end{proof}
