\documentclass{easychair}

\title{SMT-proof compression with small conflict sets}
%\titlerunning{SMT-proof compression with small conflict sets}

\author{Andreas Fellner\inst{1,2}
   \and Pascal Fontaine\inst{3}
   \and\\ Georg Hofferek\inst{4}
   \and Bruno Woltzenlogel Paleo\inst{2,5}
}
\institute{IST-Austria, Klosterneuburg (Austria)
\and Vienna University of Technology (Austria)
\and Inria, Loria, U. of Lorraine (France)
\and IAIK, Graz University of Technology (Austria)
\and Australian National University and NICTA (Australia)
}

\authorrunning{Fellner et al.}

\usepackage{amssymb}
\usepackage{graphicx}            			% Figures
\usepackage{tikz}					% tikz graphics
\usetikzlibrary{arrows,automata,positioning}
\usetikzlibrary{fit}
%\usepackage{hyperref}

\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\begin{document}

\maketitle

\begin{abstract}

This paper presents a novel proof compression algorithm for proofs in the SMT theory of equality and uninterpreted functions.
The algorithm is based on replacing congruence closure conflicts by more general, smaller ones. 
Thereby we effectively cut out literals and as a consequence whole subproofs.
However, finding the smallest conflict is NP-complete.
We propose a method to obtain short explanation as shortest paths in a weighted congruence graph.
Using this method, we show that we obtain shorter explanations than existing techniques, while remaining optimal asymptotic runtime.
The presented techniques were implemented into the proof compression tool Skeptik and evaluated on a wide range of proofs from problems of the SMT-lib.

\end{abstract}

%\bibliographystyle{plain}
%\bibliography{biblio}

\end{document}
