\documentclass[twocolumn]{svjour3}

\RequirePackage{fix-cm}

\usepackage{wrapfig}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{xcolor}
\usepackage{enumitem,amsmath,amssymb}
\usepackage{breakurl}    % used for \url and \burl
\usepackage[linesnumbered,noline,noend,vlined,ruled]{algorithm2e}
%\def\defaultHypSeparation{\hskip.1in}

\usepackage{tikz}
%\usepackage{subfig}
\usepackage{array,booktabs,multirow}
\usepackage{placeins}

\usepackage{logictools}
\usepackage{prooftheory}
\usepackage{comment}
\usepackage{mathenvironments}
\usepackage{drawproof}

\usepackage{url}

\DeclareMathOperator{\pebble}{P}
\DeclareMathOperator{\unpebble}{U}
%\DeclareMathOperator{\pebbledAt}{pebbledAt}
\DeclareMathOperator{\peb}{peb}
\DeclareMathOperator{\free}{free}
\DeclareMathOperator{\ap}{\alpha}
\DeclareMathOperator{\relPerfomance}{rp}
\newcommand{\pebbledAt}[2]{Peb^{#1}_{#2}}
\newcommand{\unpebbleAt}[2]{UPeb^{#1}_{#2}}

\newcommand{\indexIn}[3]{\ensuremath{#1 \scriptstyle \in \{#2\ldots #3\} \displaystyle}}

\usetikzlibrary{patterns}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\defeq}{\mathrel{\mathop:}=}
\newcommand{\eqdef}{\mathrel{\mathop=}:}

\newcommand{\nodedistance}{0.55cm}
\newcommand{\nodedistanceThree}{0.6cm}
\newcommand{\nodedistanceTwo}{1cm}

\newcommand{\Conclusion}[1]{\clause_{#1}}
\newcommand{\Premises}[2]{P_{#1}^{#2}}
\newcommand{\Children}[2]{C_{#1}^{#2}}
\newcommand{\Axioms}[1]{A_{#1}}

\newcommand{\axiom}[1]{\widehat{#1}}
\newcommand{\n}{v}
\newcommand{\raiz}[1]{\rho(#1)}
\newcommand{\pspace}[2]{s(#1,#2)}



\journalname{International Journal on Software Tools for Technology Transfer}

\title{Greedy Pebbling for Proof Space Compression}

\author{
  Andreas Fellner 
  \and 
  Bruno Woltzenlogel Paleo 
}

\authorrunning{A.Fellner \and B.Woltzenlogel Paleo}

\institute{
  Andreas Fellner \\
	\email{afellner@forsyte.tuwien.ac.at} \\
	Bruno Woltzenlogel Paleo  \\
	\email{bruno@logic.at} \\ \at
  Theory and Logic Group\\
	Institute for Computer Languages \\
  Vienna University of Technology
}

\begin{document}

\maketitle

\begin{abstract} %%150-250 words 
Automated reasoning tools for the verification and synthesis of software 
often produce proofs to allow independent certification of the correctness 
of the produced solutions. 
As proofs can be large, this paper 
considers the problem of compressing proofs with respect to their 
\emph{space}, which is approximately proportional to the 
memory necessary to check them. 
Proof checking with a small 
amount of available memory is analogous to playing a \emph{Pebbling Game} 
with a small number of pebbles. 
This paper exploits this 
analogy and describes novel algorithms for playing a \emph{Pebbling Game}. 
The sequence of moves executed in the pebbling game then 
corresponds to an improved topological ordering of the nodes of 
the proof, leading to smaller memory consumption when the proof 
is checked. Because the number of possible pebbling strategies and 
topological orderings is too large, brute-force approaches to find 
optimal solutions are impractical, and hence the new pebbling algorithms 
proposed here are based on heuristics for finding good, though 
not necessarily optimal, solutions. The algorithms are evaluated on the 
task of compressing the space of thousands of propositional resolution 
proofs generated by SAT- and SMT-solvers. 


\end{abstract}

%% 4-6 KEYWORDS
\keywords{Proof Compression, Memory Consumption, Resolution, Pebbling Games}

\section{Acknowledgments}

Andreas Fellner was supported by the Google Summer of Code 2013 program.
Bruno Woltzenlogel Paleo was supported by the Austrian Science Fund, project P24300.
We would like to thank Armin Biere for clarifying why resolution chains are not left- associative in the TraceCheck proof format.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Introduction}
%\label{ch:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Resolution %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Resolution}
%\label{ch:resolution}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/resolution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Pebbling Game %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Pebbling Game}
%\label{ch:pebbling-game}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_game}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Pebbling and Proof Checking %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Pebbling and Proof Processing}
%\label{ch:pebblingprocessing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\input{chapters/pebbling/pebbling_proof-processing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Pebbling as a Satisfiability Problem 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Pebbling as a Satisfiability Problem}
%\label{ch:pebblingSAT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_sat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Greedy Pebbling Algorithms %%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Greedy Pebbling Algorithms}
%\label{ch:algorithms}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_algorithms}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Heuristics %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Heuristics}
%\label{ch:heuristics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_heuristics}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Experiments %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Experiments}
%\label{ch:experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_experiments}

%\input{chapters/pebbling_futurework}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Conclusion %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\chapter{Conclusion}
%\label{ch:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/pebbling_conclusion}

%\newpage

%\marginpar{ToDo: fix overfull hboxes}

\bibliographystyle{splncs}
\bibliography{jabref_references}


\end{document}